### 2014-06-17 writeups/fb2014/abstract.tex
foo

\documentclass[english]{article}
\usepackage{graphicx}
\usepackage[T1]{fontenc}
\usepackage{babel}
\usepackage[group-separator={,}]{siunitx}
\frenchspacing
\raggedright

\title{Abstract Submission for Fourth Annual Midwestern Cognitive Science Conference}

\author{Clayton Stanley (clayton.stanley@rice.edu) \\
  Department of Psychology, 6100 Main Street \\
  Houston, TX 77005 USA 
  \and Michael D. Byrne (byrne@rice.edu) \\
  Departments of Psychology and Computer Science, 6100 Main Street \\
  Houston, TX 77005 USA \\
}

\begin{document}

\maketitle

%TC:break Abstract
\begin{abstract}
  This research explores how large-scale behavioral datasets can be used to evaluate and improve models in declarative memory.
  An author's task of choosing hashtags when composing a tweet on Twitter and choosing tags when asking a question on StackOverflow can be framed as a declarative memory retrieval problem.
  Two state-of-the-art declarative memory models were evaluated on how accurate they predict a user's chosen tags: an ACT-R inspired Bayesian model and a random permutation vector-based model.
  Experiments were designed to
  [1] discover the best way to leverage a user's prior hashtag use as a future predictor,
  [2] improve the models so that each model's strengths are incorporated into the other,
  and [3] formally evaluate and compare each model's performance when recommending hashtags on Twitter and tags on StackOverflow.
  The results show that each user's own past behavior is a strong predictor of future behavior, and past behavior can be combined with context in a straightforward way for both the Bayesian and vector-based context models.
  Also both models improve after dealing with high frequency words in some way, but they differ in how best to handle these stop words.
  This shows that there is no single best way to deal with stop words when using a co-occurrence model, and the most appropriate method is dependent on the type of co-occurrence model that is used.
  Overall, since the modifications to the two memory models were aimed at the architectural level, and these modifications improved performance,
  they may suggest task-independent general modifications that can help improve model fit to human data in a much wider range of domains.
\end{abstract}
%TC:break _main_

\end{document}