### 2014-06-17 writeups/fb2014/bio/bio.tex
foo

\documentclass[english]{article}
\usepackage{graphicx}
\usepackage[T1]{fontenc}
\usepackage{babel}
\usepackage[group-separator={,}]{siunitx}
\frenchspacing
\raggedright

\title{Bio}

\author{Clayton Stanley (clayton.stanley@rice.edu) \\
  Department of Psychology, 6100 Main Street \\
  Houston, TX 77005 USA 
}

\begin{document}

\maketitle

%TC:break Abstract
\begin{abstract}
  I am currently a PhD candidate in the Dr. Mike Byrne's Computer-Human Interaction Laboratory (CHIL) at Rice University.
  I have a mixed training background in mathematics, modeling, statistics, and human-computer interaction,
  after completing a BS in Physics at the Air Force Academy, an MA in Psychology at Rice, and 3 years in industry as a software engineer and cognitive scientist.
  Broadly I am interested in leveraging computational modeling and statistical techniques to better understand and model human behavior in human-machine systems.
  More specifically I am interested in domains in which large sets of user performance data for a particular task have been amassed.
  For my dissertation research, I am exploring how large-scale behavioral datasets like StackOverflow and Twitter can be used as a testbed to evaluate and improve models of declarative memory, and consequently, 
  improve how well the tagging models can predict each user's behavior. 
\end{abstract}
%TC:break _main_

\end{document}